Thesis: draft

### 2021-03-13 draft.tex
made some small adjustements

One of the tools to deal with these complex systems and to give us more insight in the possible underlying structures are complex networks. A network, also often called a graph, is a structure composed of nodes or vertices and a set of links or edges that indicate the interactions between the nodes (in this thesis terms like agent, actor, user, individual, people etc will be interchangebly used to denote the nodes; edges will usually just be called edges) \cite{Mata2020}.\\
Representing/modeling a complex system as a graph makes the system appear more simple and tractable, while it still includes the non-linearity of these systems. One can find the language to describe networks in mathematical graph theory. However, complex systems in real life situations often deal with a huge number of components, so the use of statistical and high-performance computing tools is inevitable \cite{Mata2020}. One could argue that the study of complex networks lies somewhere at the intersection of mathematical graph theory and statistical physics \cite{F.Costa2007}.\\

@@

Some other important concepts in network theory are the degree distribution, clustering/community structure, connectivity, etc. These concepts will be explained in depth in Chapter \ref{chap2}, Section \ref{netDef}, but let us already anticipate a bit on the case of social networks. It is found that many real-life social systems have a power law degree distribution \cite{Muchnik2013}. This heterogeneous or scale free degree distribution represents one of the three general properties of social networks. The other two are short distances and high clustering, these two are also referred to as "small world phenomenon" (note that in the literature the term "small world" sometimes only refers to the property of short distances, however, in this thesis, we will use it to indicate both a high clustering and short distances) \cite{Muchnik2013}. Ideally, our theoretical network should exhibit these three properties. Some theoretical networks, that are thoroughly studied, do not possess all of these three properties. However, they might still be used, because of their simplicity and ability to produce analytic results. When using these models one must always bear in mind their limitations to reproduce some properties encountered in real social systems. \\

@@

As stated before, the formation of an individual's opinion is the result of the interplay of many factors such as self-reflection, peer pressure, the individual's personality (eg. stubbornness), the information someone is exposed to, etc.  Opinion dynamics models should try to capture these complex processes in a simplified way. Several models have been developed ranging from simple binary models to more complex, continuous approaches \cite{Sirbu2016}.\\
The basic idea of all opinion dynamics models is that the nodes or agents in a social network have a variable that represents their opinion and that is updated according to some predefined rules. These models are obviously a simplification of real world opinion dynamics. It is however shown that they do display a lot of aspects of real opinion formation such as agreement, transitions between order (consensus) and disorder (fragmentation), polarization, formation of echo chambers (clusters of people with the same opinion), etc \cite{Sirbu2016}. \\
\newline
The opinion of the agents in the model can be either discrete or continuous. This thesis deals with models where each agent can have one of the two opinions A or B and thus only deals with the case of discrete opinions. This might come across as a huge simplification of the real world complexity of opinion formation, but also in real life people often have to choose between two competing opinions (eg. republicans or democrats, cat or dog, renting or buying a house,...). \\
\newline
The rules that determine how an agent updates his or her opinion can include many different aspects. They can, for example, be as simple as a majority rule (i.e. the majority model: if the majority of your neighbors have a certain opinion, you adopt that opinion as well) or can include a more probabilistic way of updating. A small overview of possible opinion dynamics models is given in Chapter \ref{chap3}, Section \ref{Bin} and Section \ref{Con}. \\
One can also include a resistance parameter that determines the hesitation of an agent to change to a new opinion and/or a parameter that determines the influence of an agent on others. Furthermore, the agents can have a parameter that determines whether they are active or not. The introduction of stubborn people will be discussed in Chapter \ref{chap3}, Section \ref{stubb}; the activation mechanism of nodes is discussed in Chapter \ref{chap3}, Section \ref{actMech}.\\
\newline
Since this thesis deals with the particular case of opinion dynamics on on-line social platforms, it is important to define/set up different filtering algorithms. The filtering algorithms that real social media companies use, are corporate secrets, but we know that there are three main principles of content curation: popularity, semantic and collaborative filtering \cite{Perra2019}. Popularity filtering refers to the practice of promoting content that is popular across the platform; semantic filtering means that post similar to previous consumed posts are recommended and collaborative filtering suggests posts that are similar to the ones our friends consume \cite{Perra2019}. In Chapter \ref{chap3}, Section \ref{filter} a detailed description of the filtering algorithms used in  this thesis will be given.
% add reference to chapter 2
%The rules that determine how an agent updates his or her opinion can include many different aspect, such as the higher or lower influence somebody can have on others, the resistance to change to a new opinion, peer pressure,... The structure of social networks plays an important role in aspects as peer pressure, whereas the algorithmic personalization used on social media platforms may interfere with the information someone is exposed to. Algorithmic personalization is the individualized way in which social media platforms show posts/information to each of its users.

@@

Another important tool in the study of opinion dynamics/formations in large groups of people is the use of statistical physics. This branch of physics offers a framework that relates microscopic properties of atoms, molecules,... to macroscopic observed behavior and has a wide field of applications inside and outside the world of physics.\\
The observation that large number of people display collective, `macroscopic' behavior begged for the use of the concepts and insights developed in statistical physics \cite{Sirbu2016}. The application of the theory of statistical physics on social phenomena is referred to as social physics or sociophysics. The `microscopic' constituents are now individual humans who interact with a limited number of other individuals and in that way form complex, `macroscopic' groups such as human societies. These `macroscopic' groups display stunning regularities, transitions from disorder to order, the emergence of consensus, universality, etc \cite{Buchanan2007}. The statistical physics approach of these social systems tries to explain the found regularities at large scale as collective effects of the interactions among these individuals \cite{Sirbu2016}.\\
\newline
The statistical approach of social phenomena has one big difficulty compared to the description of gases and other physical phenomena. In regular physics, the microscopic constituents are `simple' atoms or molecules, whose dynamics and interactions with each other are often well known and can be put in relatively simple laws with only a few parameters \cite{Castellano2009}. Humans, on the other hand, are far from simple entities and the dynamics of a single individual are far from known \cite{Castellano2009}. Even if the dynamics/behavior of individuals would be well know, then the interactions between them would still be complicated and it would be impossible to describe them with simple laws and only a few parameters \cite{Castellano2009}. Hence, it may seem that extracting useful and qualitative information out of social models is a nearly impossible and almost hopeless task \cite{Castellano2009}. \\
However, the use of statistical physics may also offer a way out. In most situations the qualitative (and sometimes even quantitative) properties of large scale phenomena do not depend on the microscopic details of the process \cite{Castellano2009}. Instead, the relevant aspects for the global behavior are often higher level features such as symmetries, conservation laws, dimensionality, etc \cite{Castellano2009}. This introduces a concept of universality. With this universality in mind, one can proceed to build models of human/social behavior where only the simplest and most important properties of single individuals and their interactions are taken into account \cite{Castellano2009}. An important factor herein is the comparison with real/empirical data to investigate whether or not the trends seen in real life are compatible with the outputs generated by the models \cite{Castellano2009}. Note that this interplay between theory, model, experiment and real-life data is not only an important factor in social physics, but also in regular physical research and science in general.

@@

Another concept that is closely related to the degree of a node is the notion of assortative and disassortative networks. More precisely, the notion of assortativity and disassortativity has to do with degree correlations. \\
The joint degree distribution $P(k, k')$ is the probability that an arbitrary edge connects a node with degree $k$ to a node with degree $k'$ \cite{F.Costa2007}. The conditional probability $P(k|k')$, which is another way to express dependencies between node degrees, is the probability that an arbitrary neighbor of a node with degree $k$ has a degree $k'$ \cite{F.Costa2007}
\begin{equation}
	P(k|k') = \frac{\left<k\right>P(k, k')}{kP(k)}
\end{equation}
The average degree of the nearest neighbors of nodes with degree $k$ can be computed as \cite{F.Costa2007}
\begin{equation}
	k_{nn}(k) = \sum_{k'} k'P(k'|k)
\end{equation}
If there are no correlations $k_{nn}$ is independent of $k$, and such a network is called a neutral network \cite{F.Costa2007}. If $k_{nn}(k)$ is an increasing function of $k$, then nodes with a high degree tend to connect to nodes with a high degree and the network is classified as assortative. If instead $k_{nn}(k)$ is a decreasing function of $k$, nodes with a high degree tend to connect to nodes with a low degree and the network is called disassortative \cite{F.Costa2007}.\\
However, the terms assortativity and disassortativity are also often used without taking degree correlations into account. Instead, one may look at a particular node property to define assortative and disassortative network \cite{Thed2014}. In this way, an assortative network is a network where nodes tend to connect to other nodes with similar properties as themselves, whereas in a disassortative network nodes tend to connect to nodes with different properties \cite{Thed2014}.

@@

Network communities is another feature that (complex) networks may posses. A community inside a network is loosely defined as a set of nodes that are more densely connected to nodes inside that set than to the other nodes in the network \cite{Saha2015}. More strictly, a community is defined based on two hypotheses: the connectedness hypothesis and the density hypothesis \cite{Albert2016}. In short, the connectedness hypothesis means that each member of a community should be reached through each other member of the same community \cite{Albert2016}. The density hypothesis implies that nodes inside a community are more likely to be linked to other nodes inside that community than to nodes outside the community \cite{Albert2016}. The density hypothesis narrows what could be considered a community, but it doesn't uniquely define it. Several community definitions are consistent with the density hypothesis. Let us consider three possible definitions: maximum cliques, strong community and weak community \cite{Albert2016}.\\
\newline
\textbf{Maximum cliques}\\
\newline
A clique is a complete subgraph, where a complete subgraph is defined as a set of nodes of the network where each node in the set is directly connected to all the others nodes in the same set \cite{Albert2016}. A community based on this definition would then be the largest clique in the network. This definition of community might, however, be too restrictive and, beside that, large cliques don't appear very frequent in networks \cite{Albert2016}.\\   
\newline
\textbf{Strong communities}\\
\newline
A strong community is defined such that each node inside the community has more edges to other nodes inside the same community than to nodes outside the community \cite{Albert2016}. Let us denote a community as $C$ and let $i$ be a node inside the community $C$. If we define the internal degree $k_i^{int}(C)$ as the number of edges of node $i$ with other nodes in $C$ and the external degree $k_i^{ext}(C)$ as the number of edges of node $i$ with nodes outside $C$, then we have the following condition for $C$ to be a strong community \cite{Albert2016}
\begin{equation}
	k_i^{int}(C) > k_i^{ext}(C) \qquad \forall i \in C
\end{equation}
\newline
\textbf{Weak communities}\\
\newline
A weak community is a community where the sum of the internal degree of all the nodes in the community exceeds the sum of the external degree of all the nodes in the community \cite{Albert2016}. We thus have the following condition for a weak community $C$ \cite{Albert2016}
\begin{equation}
	\sum_{i \in C} k_i^{int}(C) > \sum_{i \in C}k_i^{ext}(C)
\end{equation}

@@

Since for large $N$, $\ln{N} \ll N$, and we find that the average distance is a lot smaller than the size of the network. Generally, the logarithmic dependence of the average distance $\left<d\right>$ on the network size $N$ is one of the two defining properties of what is referred to as the "small world phenomenon"; the other property is a high clustering coefficient \cite{Easley2010}.  \\

@@

In 1998, Duncan Watts and Steve Strogatz constructed a network model that generates random networks with small world properties, such as high clustering and short average distances. Their model achieves this by interpolating between a regular ring lattice and a random network \cite{Watts1998}. The construction goes as follows \cite{Watts1998}
\begin{enumerate}
	\item Start with a regular ring lattice with $N$ nodes, each connected to $K = 2m$ neighbors ($m$ on each side)
	\item For each node $i$, rewire each edge that connects $i$ with its $K/2$ rightmost neighbors with a probability $p$. Rewiring is done in such a way that self-loops and edge duplication are avoided
\end{enumerate}
The probability $p$ allows us to tune the graph between a regular lattice ($p = 0$) and a random network ($p = 1$) \cite{Watts1998}. Usually, the following condition is required for the number of neighbors $K$: $N \gg K \gg \ln{N} \gg 1$. The first inequality assures sparse networks, whereas the second inequality guarantees that a random graph will be connected \cite{Watts1998}.\\
The regular ring lattice ($p=0$) has an average clustering coefficient $\bar{C}(0)$ and an average path length $\left<d(0)\right>$ equal to \cite{Watts1998}
\begin{align}\label{C0}
	\bar{C}(0) = \frac{3(K-2)}{4(K-1)} \approx \frac{3}{4} \ \text{for}\ K \gg 1 && \left<d(0)\right> \approx \frac{N}{2K}
\end{align}
The derivations of these formulas can be found in Appendix \ref{WS}, Section \ref{ringclus}. The regular ring lattice thus has a high clustering coefficient and an average path length that grows linearly with the system size $N$. \\
The random network ($p = 1$) on the other hand has a low clustering coefficient $\bar{C}(1)$ and a small average path length $\left<d(1)\right>$ (see Eq. (\ref{C}) and Eq. (\ref{d})) \cite{Watts1998}
\begin{align}
	\bar{C}(1) \approx \frac{K}{N} \ll 1 && \left<d(1)\right> \approx \frac{\ln{N}}{\ln{K}}
\end{align}
Watts and Strogatz then showed by numerical simulation that the average distance for the networks with $0 < p < 1$ is comparable with the average distance of a random graph, even for small $p$ \cite{Newman2000}. The clustering coefficient is approximately given by   \cite{Barrat1999}
\begin{equation}
	\bar{C} \approx \bar{C}(0)(1-p)^3
\end{equation} 
where $\bar{C}(0)$ is given by Equation (\ref{C0}). This can be seen in the following way
\begin{itemize}
	\item The probability that none of the three edges of the original triangles in the regular lattice are rewired is $(1-p)^3$
	\item The probability that edges are rewired back to each other is negligible
\end{itemize}
For rather small values of $p$, the clustering coefficient remains high. The model of Watts and Strogatz is thus able to generate random networks with a high clustering coefficient and short average distances.

>>>>

One of the tools to deal with these complex systems and to give us more insight in the possible underlying structures are complex networks. A network, also often called a graph, is a structure composed of nodes or vertices and a set of links or edges that indicate the interactions between the nodes (in this thesis terms like agent, actor, user, individual, people, etc will be interchangebly used to denote the nodes; edges will usually just be called edges) \cite{Mata2020}.\\
Representing/modeling a complex system as a graph makes the system appear more simple and tractable, while it still includes the non-linearity of these systems. One can find the language to describe networks in mathematical graph theory. However, complex systems in real-life situations often deal with a huge number of components, so the use of statistical and high-performance computing tools is inevitable \cite{Mata2020}. One could argue that the study of complex networks lies somewhere at the intersection of mathematical graph theory and statistical physics \cite{F.Costa2007}.\\

@@

Some other important concepts in network theory are the degree distribution, clustering/community structure, connectivity, etc. These concepts will be explained in depth in Chapter \ref{chap2}, Section \ref{netDef}, but let us already anticipate a bit on the case of social networks. It is found that many real-life social systems have a power law degree distribution \cite{Muchnik2013}. This heterogeneous or scale free degree distribution represents one of the three general properties of social networks. The other two are short distances and high clustering. These two are also referred to as "small world phenomenon" (note that in the literature the term "small world" sometimes only refers to the property of short distances, however, in this thesis, we will use it to indicate both a high clustering and short distances) \cite{Muchnik2013}. Ideally, our theoretical network should exhibit these three properties. Some theoretical networks, that are thoroughly studied, do not possess all of these three properties. However, they might still be used, because of their simplicity and ability to produce analytic results. When using these models one must always bear in mind their limitations to reproduce some properties encountered in real social systems. \\

@@

As stated before, the formation of an individual's opinion is the result of the interplay of many factors such as self-reflection, peer pressure, the individual's personality (eg. stubbornness), the information someone is exposed to, etc.  Opinion dynamics models should try to capture these complex processes in a simplified way. Several models have been developed ranging from simple binary models to more complex, continuous approaches \cite{Sirbu2016}.\\
The basic idea of all opinion dynamics models is that the nodes or agents in a social network have a variable that represents their opinion and that is updated according to some predefined rules. These models are obviously a simplification of real world opinion dynamics. It is however shown that they do display a lot of aspects of real opinion formation such as agreement, transitions between order (consensus) and disorder (fragmentation), polarization, formation of echo chambers (clusters of people with the same opinion), etc \cite{Sirbu2016}. \\
\newline
The opinion of the agents in the model can be either discrete or continuous. This thesis deals with models where each agent can have one of the two opinions A or B and thus only deals with the case of discrete opinions. This might come across as a huge simplification of the real world complexity of opinion formation, but also in real-life people often have to choose between two competing opinions (eg. republicans or democrats, cat or dog, renting or buying a house,...). \\
\newline
The rules that determine how an agent updates his or her opinion can include many different aspects. They can, for example, be as simple as a majority rule (i.e. the majority model: if the majority of your neighbors have a certain opinion, you adopt that opinion as well) or can include a more probabilistic way of updating. A small overview of possible opinion dynamics models is given in Chapter \ref{chap3}, Section \ref{Bin} and Section \ref{Con}. \\
One can also include a resistance parameter that determines the hesitation of an agent to change to a new opinion and/or a parameter that determines the influence of an agent on others. Furthermore, the agents can have a parameter that determines whether they are active or not. The introduction of stubborn people will be discussed in Chapter \ref{chap3}, Section \ref{stubb}; the activation mechanism of nodes is discussed in Chapter \ref{chap3}, Section \ref{actMech}.\\
\newline
Since this thesis deals with the particular case of opinion dynamics on on-line social platforms, it is important to define/set up different filtering algorithms. The filtering algorithms that real social media companies use, are corporate secrets, but we know that there are three main principles of content curation: popularity, semantic and collaborative filtering \cite{Perra2019}. Popularity filtering refers to the practice of promoting content that is popular across the platform; semantic filtering means that post similar to previous consumed posts are recommended and collaborative filtering suggests posts that are similar to the ones our friends consume \cite{Perra2019}. In Chapter \ref{chap3}, Section \ref{filter} a detailed description of the filtering algorithms used in  this thesis will be given.
% add reference to chapter 2

@@

Another important tool in the study of opinion dynamics/formations in large groups of people is the use of statistical physics. This branch of physics offers a framework that relates microscopic properties of atoms, molecules,... to macroscopic observed behavior and has a wide field of applications inside and outside the world of physics.\\
The observation that large number of people display collective, `macroscopic' behavior begged for the use of the concepts and insights developed in statistical physics \cite{Sirbu2016}. The application of the theory of statistical physics on social phenomena is referred to as social physics or sociophysics. The `microscopic' constituents are now individual humans who interact with a limited number of other individuals and in that way form complex, `macroscopic' groups such as human societies. These `macroscopic' groups display stunning regularities, transitions from disorder to order, the emergence of consensus, universality, etc \cite{Buchanan2007}. The statistical physics approach of these social systems tries to explain the found regularities at large scale as collective effects of the interactions among these individuals \cite{Sirbu2016}.\\
\newline
The statistical approach of social phenomena has one big difficulty compared to the description of gases and other physical phenomena. In regular physics, the microscopic constituents are `simple' atoms or molecules, whose dynamics and interactions with each other are often well known and can be put in relatively simple laws with only a few parameters \cite{Castellano2009}. Humans, on the other hand, are far from simple entities and the dynamics of a single individual are far from known \cite{Castellano2009}. Even if the dynamics/behavior of individuals would be well know, then the interactions between them would still be complicated and it would be impossible to describe them with simple laws and only a few parameters \cite{Castellano2009}. Hence, it may seem that extracting useful and qualitative information out of social models is a nearly impossible and almost hopeless task \cite{Castellano2009}. \\
However, the use of statistical physics may also offer a way out. In most situations the qualitative (and sometimes even quantitative) properties of large scale phenomena do not depend on the microscopic details of the process \cite{Castellano2009}. Instead, the relevant aspects for the global behavior are often higher level features such as symmetries, conservation laws, dimensionality, etc \cite{Castellano2009}. This introduces a concept of universality. With this universality in mind, one can proceed to build models of human/social behavior where only the simplest and most important properties of single individuals and their interactions are taken into account \cite{Castellano2009}. An important factor herein is the comparison with real/empirical data to investigate whether or not the trends seen in real-life are compatible with the outputs generated by the models \cite{Castellano2009}. Note that this interplay between theory, model, experiment and real-life data is not only an important factor in social physics, but also in regular physical research and science in general.

@@

Another concept that is closely related to the degree of a node is the notion of assortative and disassortative networks. More precisely, the notion of assortativity and disassortativity has to do with degree correlations. \\
The joint degree distribution $P(k, k')$ is the probability that an arbitrary edge connects a node with degree $k$ to a node with degree $k'$ \cite{F.Costa2007}. The conditional probability $P(k|k')$, which is another way to express dependencies between node degrees, is the probability that an arbitrary neighbor of a node with degree $k$ has a degree $k'$ \cite{F.Costa2007}
\begin{equation}
	P(k|k') = \frac{\left<k\right>P(k, k')}{kP(k)}
\end{equation}
The average degree of the nearest neighbors of nodes with degree $k$ can be computed as \cite{F.Costa2007}
\begin{equation}
	k_{nn}(k) = \sum_{k'} k'P(k'|k)
\end{equation}
If there are no degree correlations, $k_{nn}$ is independent of $k$. Such a network is called a neutral network \cite{F.Costa2007}. If $k_{nn}(k)$ is an increasing function of $k$, then nodes with a high degree tend to connect to nodes with a high degree and the network is classified as assortative. If instead $k_{nn}(k)$ is a decreasing function of $k$, nodes with a high degree tend to connect to nodes with a low degree and the network is called disassortative \cite{F.Costa2007}.\\
However, the terms assortativity and disassortativity are also often used without taking degree correlations into account. Instead, one may look at a particular node property to define assortative and disassortative network \cite{Thed2014}. In this way, an assortative network is a network where nodes tend to connect to other nodes with similar properties as themselves, whereas in a disassortative network nodes tend to connect to nodes with different properties \cite{Thed2014}.

@@

Network communities are another feature that (complex) networks may posses. A community inside a network is loosely defined as a set of nodes that are more densely connected to nodes inside that set than to the other nodes in the network \cite{Saha2015}. More strictly, a community is defined based on two hypotheses: the connectedness hypothesis and the density hypothesis \cite{Albert2016}. In short, the connectedness hypothesis means that each member of a community should be reached through each other member of the same community \cite{Albert2016}. The density hypothesis implies that nodes inside a community are more likely to be linked to other nodes inside that community than to nodes outside the community \cite{Albert2016}. The density hypothesis narrows what could be considered a community, but it doesn't uniquely define it. Several community definitions are consistent with the density hypothesis. Let us consider three possible definitions: maximum cliques, strong community and weak community \cite{Albert2016}.\\
\newline
\textbf{Maximum cliques}\\
\newline
A clique is a complete subgraph, where a complete subgraph is defined as a set of nodes of the network where each node in the set is directly connected to all the others nodes in the same set \cite{Albert2016}. A community based on this definition would then be the largest clique in the network. This definition of community might, however, be too restrictive and, beside that, large cliques don't appear very frequent in networks \cite{Albert2016}.\\   
\newline
\textbf{Strong communities}\\
\newline
A strong community is defined such that each node inside the community has more edges to other nodes inside the same community than to nodes outside the community \cite{Albert2016}. Let us denote a community as $C$ and let $i$ be a node inside the community $C$. If we define the internal degree $k_i^{int}(C)$ as the number of edges that connect node $i$ with other nodes in $C$ and the external degree $k_i^{ext}(C)$ as the number of edges that connect node $i$ with nodes outside $C$, then we have the following condition for $C$ to be a strong community \cite{Albert2016}
\begin{equation}
	k_i^{int}(C) > k_i^{ext}(C) \qquad \forall i \in C
\end{equation}
\newline
\textbf{Weak communities}\\
\newline
A weak community is a community where the sum of the internal degree of all the nodes in the community exceeds the sum of the external degree of all the nodes in the community \cite{Albert2016}. We thus have the following condition for a weak community $C$ \cite{Albert2016}
\begin{equation}
	\sum_{i \in C} k_i^{int}(C) > \sum_{i \in C}k_i^{ext}(C)
\end{equation}

@@

Since for large $N$, $\ln{N} \ll N$, we find that the average distance is a lot smaller than the size of the network. Generally, the logarithmic dependence of the average distance $\left<d\right>$ on the network size $N$ is one of the two defining properties of what is referred to as the "small world phenomenon"; the other property is a high clustering coefficient \cite{Easley2010}.  \\

@@

In 1998, Duncan Watts and Steve Strogatz constructed a network model that generates random networks with small world properties, such as high clustering and short average distances. Their model achieves this by interpolating between a regular ring lattice and a random network \cite{Watts1998}. The construction goes as follows \cite{Watts1998}
\begin{enumerate}
	\item Start with a regular ring lattice with $N$ nodes, each connected to $K = 2m$ neighbors ($m$ on each side)
	\item For each node $i$, rewire each edge that connects $i$ with its $K/2$ rightmost neighbors with a probability $p$. Rewiring is done in such a way that self-loops and edge duplication are avoided
\end{enumerate}
The probability $p$ allows us to tune the graph between a regular lattice ($p = 0$) and a random network ($p = 1$) \cite{Watts1998}. Usually, the following condition is required for the number of neighbors $K$: $N \gg K \gg \ln{N} \gg 1$. The first inequality assures sparse networks, whereas the second inequality guarantees that a random graph will be connected (thus the network is not so sparse that it becomes disconnected) \cite{Watts1998}.\\
The regular ring lattice ($p=0$) has an average clustering coefficient $\bar{C}(0)$ and an average path length $\left<d(0)\right>$ equal to \cite{Watts1998}
\begin{align}\label{C0}
	\bar{C}(0) = \frac{3(K-2)}{4(K-1)} \approx \frac{3}{4} \ \text{for}\ K \gg 1 && \left<d(0)\right> \approx \frac{N}{2K}
\end{align}
The derivations of these formulas can be found in Appendix \ref{WS}, Section \ref{ringclus}. The regular ring lattice thus has a high clustering coefficient and an average path length that grows linearly with the system size $N$. \\
The random network ($p = 1$) on the other hand has a low clustering coefficient $\bar{C}(1)$ and a small average path length $\left<d(1)\right>$ (see Eq. (\ref{C}) and Eq. (\ref{d})) \cite{Watts1998}
\begin{align}
	\bar{C}(1) \approx \frac{K}{N} \ll 1 && \left<d(1)\right> \approx \frac{\ln{N}}{\ln{K}}
\end{align}
Watts and Strogatz then showed by numerical simulation that the average distance for the networks with $0 < p < 1$ is comparable with the average distance of a random graph, even for small $p$ \cite{Newman2000}. The clustering coefficient is approximately given by   \cite{Barrat1999}
\begin{equation}
	\bar{C} \approx \bar{C}(0)(1-p)^3
\end{equation} 
where $\bar{C}(0)$ is given by Equation (\ref{C0}). This can be seen in the following way
\begin{itemize}
	\item The probability that none of the three edges of the original triangles in the regular lattice are rewired is $(1-p)^3$
	\item The probability that edges are rewired back to each other is negligible
\end{itemize}
For rather small values of $p$, the clustering coefficient remains high. The model of Watts and Strogatz is thus able to generate random networks with a high clustering coefficient and short average distances.
